---
title: "Homework 1"
author: "Natasha Woon"
date: "2026-09-25"
output: github_document
---

# Problem 1 

## Load Packages

```{r}
library(tidyverse)
```

```{r}
data("penguins", package = "palmerpenguins")
```

## Description of Penguins Data set 
The `penguins` data set contains data on penguins from three different species (Adelie, Chinstrap and Gentoo). Important variables include species, island, bill length, bill depth, flipper length, body mass, sex, and year.

This data set has `r nrow (penguins)` observations and `r ncol(penguins)` columns.

The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.

## Scatterplot 

```{r warning = F}
ggplot(data = penguins,
       mapping = aes(bill_length_mm, flipper_length_mm, color = species)) + geom_point()

ggsave("scatterplot.png")
```

# Problem 2

```{r}
set.seed(123)

problem2_df = 
  tibble(
    random_sample = rnorm(10),
    logical_vector = (random_sample > 0),
    character_vector = c("Dog", "Giraffe", "Hamster", "Frog", "Elephant", "Cat", "Moose", "Otter", "Lizard", "Kangaroo"), 
    factor_vector = factor (c("Medium", "Large", "Small", "Small", "Large", "Medium", "Large", "Medium", "Small", "Large"))
  )

problem2_df

mean(pull(problem2_df, random_sample))

mean(pull(problem2_df, logical_vector))

mean(pull(problem2_df, character_vector))

mean(pull(problem2_df, factor_vector))

```

After taking the mean of each variable in the dataset, it is clear that the mean value can only be generated for numerical and logical vectors. Numerical variables can be averaged. Logical variables treat TRUE as 1 and FALSE and 0, and can therefore produce and average. However, character and factor vectors produced a warning NA result, and were not able to generate a mean value when utilizing the mean() function. This makes sense because R cannot generate a mean of values such as `dog` and and `lizard`.

```{r, results = "hide"}
as.numeric(pull(problem2_df, logical_vector))
as.numeric(pull(problem2_df, character_vector))
as.numeric(pull(problem2_df, factor_vector))
```

Applying the `as.numeric` function to the logical vector works since TRUE or FALSE can be converted to 1 or 0s, and thus, a mean can be taken. However, the `as.numeric` function cannot be applied to character vectors because the character strings cannot be converted into numbers. Lastly, the `as.numeric` function converts the factor variables into the number assigned to each factor level, and does not actually convey the factors (Small, Medium or Large) into numbers, and therefore taking the mean here would not be meaningful.

